---
title: "Modeling in tidyverse"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  html_notebook: default
date: "18 February 2018"
---

Most recent addition to tidyverse is `rsample` which introduces new way of peformning advanced resampling of the data. Package documentation is [here](https://topepo.github.io/rsample). A little older but incredibly useful package is `modelr`which has a lot of helper functions for evaluating models. We will only scratch a surface today with simple linear models, mostly focusing on different approaches to cross-validation.

```{r}
library(tidyverse)
library(rsample)
library(broom)
library(modelr)

gapminder <- gapminder::gapminder

models_df <- tribble(
  ~mdl,        ~frml,
  "yr",         as.formula(lifeExp ~ year),
  "gdp",        as.formula(lifeExp ~ log(gdpPercap)),
  "yr_gdp",     as.formula(lifeExp ~ year + log(gdpPercap)),
  "yr_pop_gdp", as.formula(lifeExp ~ year + pop + log(gdpPercap))
)

```


```{r}
cv_samples <- gapminder %>% 
  group_by(continent) %>% nest() %>% 
  mutate(cv=map(data, vfold_cv)) %>% 
  unnest(cv)
cv_samples

cv_models_df <- cv_samples %>% 
  crossing(models_df)

perf_metrics <- cv_models_df %>% 
  mutate(fits=map2(frml, splits, ~lm(formula=.x, data=analysis(.y))),
         rmse=map2_dbl(fits, splits, ~modelr::rmse(.x, data=assessment(.y))))

ggplot(perf_metrics)+
  geom_boxplot(aes(x=continent, y=rmse))+
  facet_wrap( ~mdl)
```

It seems like more complex model wins. But is it really going to help us in predicting the future?

Lets do time-aware resampling

```{r}

cv_samples <- gapminder %>% 
  group_by(continent, country) %>% nest() %>% 
  mutate(ro_rsmpl=map(data, rolling_origin, initial=7)) %>% 
  unnest(ro_rsmpl)

cv_models_df <- cv_samples %>% 
  tidyr::crossing(models_df)

resid_df <- cv_models_df %>% 
  mutate(fits = map2(frml, splits, ~lm(formula=.x, data=analysis(.y))),
         rmse = map2_dbl(fits, splits, ~modelr::rmse(.x, data=assessment(.y)))) 

ggplot(resid_df) + 
  geom_line(aes(x=id, y=rmse, group=country, color=continent))+
  facet_wrap(~mdl)


resid_df %>% filter(mdl=="yr",rmse>10) %>% arrange(continent, country, id)
```

